---
title: "Motor Trend"
author: "Chandrasekar Ganesan"
date: "March 27, 2017"
output:
  pdf_document: default
---

#Instructions

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

* “Is an automatic or manual transmission better for MPG”
* "Quantify the MPG difference between automatic and manual transmissions"


#Load the Data

```{r}
data(mtcars)
summary(mtcars)

```

#Exploratory Analysis

```{r}
temp <- mtcars
temp$am <- as.factor(temp$am)
levels(temp$am) <- c("Automatic","Manual")
plot(temp$am, temp$mpg, ylab="Miles Per Gallon (mpg)",
     main="Effect of Transmission to MPG")

mauto <- median(temp[temp$am=="Automatic",]$mpg)
mman <- median(temp[temp$am=="Manual",]$mpg)

arrows(1.5,mauto, 1.5, mman, lty=1, lwd=2, col="red", length=0.25, angle=20)


```

As it can be clearly seen, that **mgp** is higher in vehicles with manual transmission when compared to vehicles with Automatic transmission.

#Deeper Dive

#Hypothesis

Let's define a two sided hypothesis test as follows:

1) **H0** - Miles per gallon is not influenced by automatic or manual transmission (null hypothesis)
2) **H1** - Miles per gallon is influenced by automatic or manual transmission

Let's quantify the variation via regression analysis.

Understanding the correlations of mtcars dataset

```{r}
data(mtcars)
library(corrgram)
corrgram(mtcars, order=TRUE, lower.panel=panel.shade,
   upper.panel=panel.pie, text.panel=panel.txt,
   main="Car Milage Data in PC2/PC1 Order")
```

It is evident that **mpg** is highly correlated with **gear, am, wt, disp, cyl, hp** and **qsec**

But it is imperative we reduce the variable set to avoid overfitting or underfitting.

Let's use elimination process to remove unrelated variables. Do a multi variable regression with mpg as outcome based on all variables.

```{r}
fit <- lm(mpg ~ ., data=mtcars)
summary(fit)$coef
```

Looking at the result **cyl** has a high P value of **0.91608738**. Let's eliminate this and rerun the regression with the rest as follows:

```{r}
fit <- lm(mpg ~ disp+hp+drat+wt+qsec+vs+am+gear+carb, data=mtcars)
summary(fit)$coef
```

We repeat the process recursively until we get a model with dependent variables that has P value < 0.05

The model variables that impacts **mpg** are **wt, qsec, am** as shown below. Let's run the final regression on the selected dependent variables.

```{r}
fit <- lm(mpg ~ wt+qsec+am -1, data=mtcars)
summary(fit)$coef
```

Observe that P-value for **am** is 2.329423e-04, much less than 0.05. We have enough evidence to reject the null hypothesis *(H0)*

#Exploring Results

```{r}
par(mfrow=c(2,2), mar=c(2,2,2,.5), font.main=1, font.sub=1, cex.lab=1, cex.axis=1)
plot(fit, cex=0.8, pch=19, col=rgb(0,1,.75,0.4))
```

#Conclusion

Our plot is not showing any pattern. The linear model we have is a reasonable fit.

Let's quantify the MPG difference between *automatic* and *manual transmission*

```{r}
coefs <- summary(fit)$coef
mpgRange <- coefs["am",1] + c(-1,1) * qt(0.975, df = fit$df) * coefs["am",2]
print(mpgRange)

```

Based on 95% confidence, we can estimate that *switching to manual transmission from automatic transmission will result in an average of 2.2 to 6.4 miles per gallon*.

Hence manual transmission is **better** than automatic transmission



